Wireless Automobile Detection, License Plate Processing, and Data Availability Network

### 2014-05-01 paper/paper.tex

\begin{abstract}
{\color{red}\textbf{NOT COMPLETE}}
This paper describes a parking lot monitoring system to be used at the Colorado School of Mines (CSM). The system will help campus members enjoy a better parking experience by spending less time trying to find open lots. This is a semester project for the Computer Science graduate course \emph{Wireless Sensor Networks}, taught by Dr. Tracy Camp at CSM. The work was an extension of a project started by campus computing group ACMx, which uses a wireless sensor network to track the number of cars in specific parking lots. The system will extend the ACMx project by adding components to the sensor network that record the license plate numbers of cars as they enter and exit lots. The information gathered by the sensor network will be displayed on a website that will allow users to survey the state of the lots before arriving, allowing them to make informed parking decisions. The website will also be the result of collaboration between the authors of this proposal and the ACMx group. The proposal provides some background information, describes the proposed system and its components, and discusses implementation strategies. The goal of the project is a real-world deployment on the CSM campus by May 2014.
\end{abstract}

@@

The second part of the project is motivated by a desire to test the feasibility of using the small Raspberry Pi Linux computer as a platform for capturing and processing images of license plates. Because the Raspberry Pi is inexpensive (\$25-\$35 depending on the model), it could be useful in a wide variety of applications that require accurate license plate recognition. Currently, our web application simply lists the license plate numbers for site administrators to view, but we envision that this monitoring system may eventually be used by the campus parking department to detect when vehicles have entered lots without parking passes. The system itself will consist of several subsystems, which are described in detail in section 2.

@@

\begin{table}[h!]
\begin{center}
\begin{tabular}{| l || l | l | l |}
\hline
Data Set & Correct Recognition & Total Images & Percent Correct \\ \hline
Unaltered images & 20 & 52 & 38.46 \% \\ \hline
Cropped images & 4 & 52 & 7.69 \% \\ \hline
Compressed images & 4 & 27 & 14.81 \% \\ \hline
\end{tabular}
\end{center}
\caption{OpenALPR Results}
\label{tab:openalprresults}
\end{table}

\begin{table}[h!]
\begin{center}
\begin{tabular}{| l || l | l | l |}
\hline
Data Set & Correct Recognition & Total Images & Percent Correct \\ \hline
Unaltered images & 34 & 52 & 65.39 \% \\ \hline
Cropped images & 12 & 52 & 23.08 \% \\ \hline
Compressed images & 8 & 27 & 29.63 \% \\ \hline
\end{tabular}
\end{center}
\caption{Intrada ALPR Results}
\label{tab:intradaresults}
\end{table}
Three data sets were run through each ALPR system. As seen in Table \ref{tab:openalprresults} and Table \ref{tab:intradaresults}, Intrada ALPR outperformed OpenALPR in each case in terms of accuracy. In terms of speed, OpenALPR always outperformed Intrada ALPR by at least a factor of three. In some cases, OpenALPR was better by a factor of 110.

\begin{figure}
\begin{center}
\includegraphics[width=4.5in]{unexpectedrecognition}
\end{center}
\caption{The foreground license plate goes unrecognized in favor of a smaller plate in the background}
\label{fig:unexpectedrecognition}
\end{figure}

These results were calculated by determining whether or not the ALPR software correctly identified \emph{any} license plate in the photo. It is the case that, for certain out of state plates or nonstandard plates (e.g. Disabled Veteran), the ALPR programs will not identify the primary, or largest, license plate in the photo. However, both programs were successful in recognizing a much smaller, but standard, license plate. Figure \ref{fig:unexpectedrecognition} shows an example image where a New Mexico plate is ignored in favor of a smaller plate in the background.
\section{Future Research}
There is substantial future research that can be performed to add features and improve the efficacy of the system. This area of research appears very promising as similar systems may be applied and adapted to many different settings.

\subsection{ALPR Training}
The accuracy rate for our ALPR software is not as high as we had originally hoped it to be. OpenALPR offers a separate program, 'train-ocr' to help build definition files for certain geographic regions. Because it initially came seeded with training data for the entire continent, the recognition of Colorado plates may be improved if we create a training file for only Colorado.

Intrada ALPR cannot be configured by the end user (our team). We can, however, send sample images from our deployment to Q-Free, where our files can be used to train our account for better recognition.

Because we ended up using two ALPR systems instead of one, it would be beneficial to utilize both to provide a confidence level for our project. Having validation from both ALPRs would likely mean the characters that were extracted were correct. Extractions that were not correct could be flagged in the database for human intervention.

A proposed long-term goal for this project was to integrate license recognition with CSM's parking database to identify users who do not have valid parking permits. With the work that has been completed, we are now at a place where we could begin coding an interface to pass license data to Parking and Facilities.

@@

@@

\section{Summary}
{\color{red}\textbf{NOT COMPLETE}}
This project extends the work done by ACMx to create the first comprehensive parking management system for the Colorado School of Mines. The system's expansions cover interfacing with the existing automobile detection network, acquiring images of license plates, processing said images, and presenting the resulting data in a web application.

@@

Image acquisition will come from several acquisition Raspberry Pis stationed at parking lot entrances. These images will be sent to a central base station, which will then upload the data to a server. Once on the server, additional image processing will be done to retrieve the characters composing the license plate. This data will be stored in a database for use in the web application.

In addition, the base station Raspberry Pi will sample images of the parking lots to allow administrators to correct the status of the system. This may happen when the acquisition motes unsuccessfully identify a car entering or exiting, and therefore have incorrect information regarding the parking lot's capacity.

The web application will serve to show the state of the parking lots, as well as offer a way for administrators to reconcile observed data (from base station images) and sensed data (from acquisition motes). Administrators will also be able to access license plate data from processed images to show proof-of-concept for this utility.

Once deployed, the SmartLots system, augmented with our additional license plate extraction subsystem and improvements to the web application, will offer the Mines campus an accessible way to manage campus parking lots. Faculty and students will be able to use this to find open parking and save time. Future applications could include interfacing with Parking Services to locate illegally parked vehicles or to gather data regarding campus parking needs.

>>>>

\begin{abstract}
This paper describes a parking lot monitoring system to be used at the Colorado School of Mines (CSM). The system is designed to help campus members enjoy a better parking experience by spending less time trying to find open parking lots. This project was a semester project for the computer science graduate course \emph{Wireless Sensor Networks}, taught by Dr. Tracy Camp at CSM. The work described in this paper is an extension of a project called Smartlots started by campus computing group ACMx, which uses a wireless sensor network to track the number of cars in specific parking lots. The system extends the ACMx project by adding components to the sensor network that record the license plate numbers of cars as they enter and exit lots, and by adding an administrator interface to the already existing Smartlots website. The information gathered by the sensor network is displayed on a website that lets users survey the state of the lots before arriving on campus, allowing them to make informed parking decisions. The website is also the result of collaboration between the authors of this paper and the ACMx group. The paper provides some background information, describes the parking lot monitoring system and its components, discusses the details of our implementation, and provides some initial results from testing the system and its various components.
\end{abstract}

@@

The second part of the project is motivated by a desire to test the feasibility of using the small Raspberry Pi Linux computer as a platform for capturing and processing images of license plates. Because the Raspberry Pi is inexpensive (\$25-\$35 depending on the model), it could be useful in a wide variety of applications that require accurate license plate recognition. Currently, our web application simply lists the license plate numbers for site administrators to view, but we envision that this monitoring system may eventually be used by the campus parking department to detect when vehicles have entered lots without parking passes. The system itself consists of several subsystems, which are described in detail in section 2.

@@

\begin{table}[h!]
\begin{center}
\begin{tabular}{| l || l | l | l |}
\hline
Data Set & Correct Recognition & Total Images & Percent Correct \\ \hline
Unaltered images & 12 & 41 & 29.27 \% \\ \hline
Cropped images & 4 & 52 & 7.69 \% \\ \hline
Compressed images & 4 & 27 & 14.81 \% \\ \hline
\end{tabular}
\end{center}
\caption{OpenALPR Results}
\end{table}

\begin{table}[h!]
\begin{center}
\begin{tabular}{| l || l | l | l |}
\hline
Data Set & Correct Recognition & Total Images & Percent Correct \\ \hline
Unaltered images & 27 & 41 & 65.85 \% \\ \hline
Cropped images & 12 & 52 & 23.08 \% \\ \hline
Compressed images & 8 & 27 & 29.63 \% \\ \hline
\end{tabular}
\end{center}
\caption{Intrada ALPR Results}
\end{table}

\section{Future Research}
There is substantial future research that can be performed to add features and improve the efficacy of the system. This area of research appears very promising as similar systems may be applied and adapted to many different settings.

@@

\subsection{Real-world Deployment}
Due to the fact that this project was time-constrained, we were unable to test our system in a real-world environment. We had initially planned to deploy a working version of our system on two parking lots on the CSM campus. Even though we never deployed our system, we were able to get everything connected together as described in section 2, so that data collected by the sensors flowed all the way up through the system and was displayed in the web application. Our plans for a future deployment are as follows: ACMx and members from our group plan to deploy a version of our system at the CTLM Upper and Lower parking lots at CSM. These lots were chosen because of their accessible location, unobstructed visibility from a campus structure, and frequent use. Both parking lots have one entryway, each of which will be equipped with an automobile detection subsystem. The junction between the two lots is also of interest, because we want to know when a car moves from the upper lot to the lower lot, and vice versa. The locations of the entrances and junction are shown in Figure \ref{fig:deployment}. The locations for the three sensor nodes benefit from having a clear view of the sky for increased photovoltaic efficiency.
\begin{figure}
\begin{center}
\includegraphics[width=5in]{deployment}
\end{center}
\caption{Deployment Location}
\label{fig:deployment}
\end{figure}

The central base station, as described in section 2.4, will be placed in an upper window of the CTLM building, allowing it to take a clear overhead picture of at least one of the two parking lots. The first deployment only have one overhead camera to start off with, although ideally we would have two - one for each lot. Several potential locations exist which allow for a clear line-of-sight to all automobile detection subsystems, one of which is detailed in Figure \ref{fig:deployment}. The location of the base station must put it in range of the XBees mounted on each sensor node, to allow data to be transferred from the detection subsystem to the base station. The locations being considered must also provide internet access to the base station Raspberry Pi through an Ethernet port.

\section{Summary}
This project extends the work done by ACMx towards creating the first comprehensive, computerized parking management system for the Colorado School of Mines. These extensions cover interfacing with the previously existing automobile detection network, acquiring images of license plates, processing those images, and presenting the resulting data in a web application.

@@

\section{Summary}

This project extends the work done by ACMx to create the first comprehensive parking management system for the Colorado School of Mines. The system's expansions cover interfacing with the existing automobile detection network, acquiring images of license plates, processing said images, and presenting the resulting data in a web application.

@@

Image acquisition is done by Raspberry Pis stationed at parking lot entrances. These images are sent to a central base station Raspberry Pi, which then uploads the data to the server via our RESTful API. Once on the server, additional image processing is done to retrieve the characters composing the license plate. This data is then stored in a database for use in the web application.

Acquiring images of license plates includes listening for an interrupt from the automobile detection sensor, temporarily storing the image, and on-mote pre-processing of the image. By completing some processing of the image on the Raspberry Pi, we reduce the size of data we need to transmit, which saves battery power while also reducing the transmission time, thereby improving system latency and throughput.

The web application serves to show the state of the parking lots, and provides an interface that allows administrators to view license plate data and sensor data, and manage user accounts.

Once deployed, the Smartlots system, augmented with our additional license plate extraction subsystem and improvements to the web application, will offer the Mines campus an accessible way to manage campus parking lots. Faculty and students will be able to use this to find open parking and save time. Future applications could include interfacing with Parking Services to locate illegally parked vehicles or to gather data regarding campus parking needs.
